---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# ofpetrial: Design On-farm Precision Experiments 

<!-- badges: start -->
<!-- badges: end -->

The `ofpetrial` package allows the user to design agronomic input experiments in a reproducible manner without using ArcGIS or QGIS.  

## Installation

You can install the development version of ofpetrial from [Github](https://github.com/DIFM-Brain/TrialDesign):

```{r, eval = FALSE}
devtools::install_github("DIFM-Brain/ofpetrial")
```

## Example

Here, we demonstrate how to use the `ofpetrial` package to create two-input on-farm experiment trial designs.

### Create experimental plots

We start with specifying plot and machine information for inputs using `make_input_plot_data()`, which simply creates a data.frame of the specified information with some interal unit conversion of length (feet to meter).

```{r}
seed_plot_info <-
  make_input_plot_data(
    form = "seed",
    plot_width = 30,
    machine_width = 60,
    section_num = 24,
    length_unit = "feet"
  )

seed_plot_info

n_plot_info <-
  make_input_plot_data(
    form = "NH3",
    plot_width = measurements::conv_unit(60, "ft", "m"),
    machine_width = measurements::conv_unit(60, "ft", "m"),
    section_num = 1
  )

n_plot_info
```

Now that plot and machine specifications for the inputs are ready, we can create experiment plots based on them using `make_exp_plots()`. 

```{r}
input_plot_info <- list(seed_plot_info, n_plot_info)

exp_data <-
  make_exp_plots(
    input_plot_info = input_plot_info,
    boundary_file = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_file = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    harvester_width = 30,
    abline_type = "free",
    headland_length = 30,
    side_length = 60,
    min_plot_length = 200,
    max_plot_length = 300,
    length_unit = "feet",
    perpendicular = FALSE
  )

exp_data$exp_plots
```

We can visualize the layout of the experiment plots using `viz_ep()`.

```{r, dpi = 600}
viz_ep(exp_data)
```

### Assign rates

Let's now assign input rates to the experimental plots we just created. Before doing so, we need to prepare rate information for both inputs using `make_input_rate_data()`.

```{r}
#!===========================================================
#! Assign rates
#!===========================================================
seed_rate_info <-
  make_input_rate_data(
    plot_info = seed_plot_info,
    gc_rate = 32000,
    unit = "seed",
    min_rate = 16000,
    max_rate = 40000,
    num_rates = 5,
    design_type = "jcls"
  )

seed_rate_info

n_rate_info <-
  make_input_rate_data(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "jcls",
  )

n_rate_info
```

We can now use `assign_rates()` to assign rates to experiment plots.

```{r}
trial_design <- assign_rates(exp_data, list(seed_rate_info, n_rate_info))
```

Here is the visualization of the trial design done by `viz_td()`.

```{r, dpi = 600}
viz_td(trial_design)
```
