---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ofpetrial: Design On-farm Precision Experiments

<!-- badges: start -->
[![R-CMD-check](https://github.com/DIFM-Brain/ofpetrial/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/DIFM-Brain/ofpetrial/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `ofpetrial` package allows the user to design agronomic input experiments in a reproducible manner without using ArcGIS or QGIS. The vignette for this package is [here](https://difm-brain.github.io/ofpetrial/).

## Installation

You can install the development version of ofpetrial from [Github](https://github.com/DIFM-Brain/TrialDesign):

```{r, eval = FALSE}
devtools::install_github("DIFM-Brain/ofpetrial")
```

## Example

Here, we demonstrate how to use the `ofpetrial` package to create two-input on-farm experiment trial designs.

```{r}
library(ofpetrial)
```

### Create experimental plots

We start with specifying plot and machine information for inputs using `prep_plot`, which simply creates a data.frame of the specified information with some internal unit conversion of length (feet to meter).

```{r}
seed_plot_info <-
  prep_plot(
    input_name = "seed",
    unit_system = "imperial",
    machine_width = 60,
    section_num = 24,
    harvester_width = 30,
    plot_width = 30
  )

seed_plot_info

n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 30,
    section_num = 1,
    harvester_width = 30,
    plot_width = 30
  )

n_plot_info
```

Now that plot and machine specifications for the inputs are ready, we can create experiment plots based on them using `make_exp_plots()`. 

```{r}
input_plot_info <- list(seed_plot_info, n_plot_info)

exp_data <-
  make_exp_plots(
    input_plot_info = input_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )
```

The experiment plots created by `make_exp_plots()` is stored in `exp_plots`.

```{r}
exp_data$exp_plots
```

`exp_data$exp_plots` is a list and you can access the individual experiment plots (an `sf` object) like this.

```{r}
exp_data$exp_plots[[1]]
```

We can visualize the layout of the experiment plots using `viz_layout()`.

```{r, dpi = 600}
viz(exp_data, type = "layout", abline = TRUE)
```

### Assign rates

Let's now assign input rates to the experimental plots we just created. Before doing so, we need to prepare rate information for both inputs using `prep_rates()`.

```{r}
#!===========================================================
# ! Assign rates
# !===========================================================
seed_rate_info <-
  prep_rate(
    plot_info = seed_plot_info,
    gc_rate = 32000,
    unit = "seed",
    min_rate = 16000,
    max_rate = 40000,
    num_rates = 5,
    design_type = "jcls"
  )

seed_rate_info

n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "ls",
    rank_seq_ws = c(5, 4, 3, 2, 1)
  )

n_rate_info
```

We can now use `assign_rates()` to assign rates to experiment plots (see [this vignette](https://difm-brain.github.io/ofpetrial/articles/V1-trial-design-options.html) for other design options and [this vignette](https://difm-brain.github.io/ofpetrial/articles/V3-change-rates-manually.html) for changing rates manually.).

```{r}
trial_design <- assign_rates(exp_sf = exp_data, rate_info = list(seed_rate_info, n_rate_info))
```

Here is the visualization of the trial design done by `viz`.

```{r, dpi = 600}
viz(trial_design)
```

Along with the spatial pattern of the input rates, the applicator/planter ab-line and harvester ab-line are drawn by default.

### Diagnose the trial design

The `ofpetrial` package offers several functions to check the soundness of a trial design (see [this vignette](https://difm-brain.github.io/ofpetrial/articles/V4-diagnose-td.html) for more details and examples). 

Here, let's check the correlation between the seed and NH3 rates. 

```{r}
(
  cor_inputs <- check_ortho_inputs(trial_design)
)
```

The correlation coefficient is `r round(cor_inputs, digits = 2)`.

### Write the trial design files for implementation

Once you are satisfied with the trial design adn the location of the ab-lines, you can write out all the necessary files as shape files for the machine operators to actually implement the trial.

```{r, eval = FALSE}
write_trial_files(td, zip = TRUE, zip_name = "td-collection")
```

Here, all the shape files (trial dsign, applicator/planter ab-line, and harvester ab-line) are packed in a zipped folder, named "td-collection.zip".

# Acknowledgement

This project was funded in part by a United States Department of Agriculture—National Institute of Food and Agriculture (USDA—NIFA) Food Security Program Grant (Award Number 2016-68004-24769) and by United States Department of Agriculture (USDA) -Natural Resources Conservation Service (NRCS), Commodity Credit Corporation (CCC), Conservation Innovation Grants On-Farm Conservation Innovation Trials (Award Number USDA-NRCS-NHQ-CIGOFT-20-GEN0010750).
