---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300
)
```

# ofpetrial: Design on-Farm Precision Experiments

<!-- badges: start -->
[![R-CMD-check](https://github.com/DIFM-Brain/ofpetrial/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/DIFM-Brain/ofpetrial/actions/workflows/R-CMD-check.yaml)
`r badger::badge_cran_checks("ofpetrial")`
`r badger::badge_repostatus("Active")`
`r badger::badge_cran_download("ofpetrial", "grand-total", "blue")`
<!-- badges: end -->

<!-- <a href = "https://difm-brain.github.io/ofpetrial/" target = "_blank"><img src="https://cranlogs.r-pkg.org/badges/grand-total/ofpetrial"></a> -->

The `ofpetrial` package allows the user to design agronomic input experiments in a reproducible manner without using ArcGIS or QGIS. The [vignette](https://difm-brain.github.io/ofpetrial/) for this package provides more detailed guidance on how to use the package.

## Installation

### CRAN version

You can install the CRAN version of the `ofpetrial` package.

```{r, eval = FALSE}
install.packages("ofpetrial")
```

### Development version

You can install the development version of ofpetrial from [Github](https://github.com/DIFM-Brain/ofpetrial):

```{r, eval = FALSE}
devtools::install_github("DIFM-Brain/ofpetrial")
```

## Example

Here, we demonstrate how to use the `ofpetrial` package to create a single-input on-farm experiment trial design (more detailed instructions on the basic workflow is provided in [this article](https://difm-brain.github.io/ofpetrial/articles/V0-basic-work-flow.html)).  

```{r}
library(ofpetrial)
```

### Create experimental plots

We start with specifying plot and machine information for inputs using `prep_plot`.

```{r}
n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 30,
    section_num = 1,
    harvester_width = 30,
    plot_width = 30
  )
```

Now, we can create experiment plots based on them using `make_exp_plots()`. 

```{r}
exp_data <-
  make_exp_plots(
    input_plot_info = n_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )

viz(exp_data, type = "layout", abline = TRUE)
```

### Assign rates

We first prepare nitrogen rates.

```{r}
#!===========================================================
# ! Assign rates
# !===========================================================
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "ls",
    rank_seq_ws = c(5, 4, 3, 2, 1)
  )

dplyr::glimpse(n_rate_info)
```

We can now use `assign_rates()` to assign rates to experiment plots.

```{r}
trial_design <- assign_rates(exp_data, rate_info = n_rate_info)
```


Here is the visualization of the trial design done by `viz`.

```{r}
viz(trial_design)
```

Along with the spatial pattern of the input rates, the applicator/planter ab-line and harvester ab-line are drawn by default.

### Write the trial design files for implementation

You can write out the trial design as a shape file.

```{r, eval = FALSE}
write_trial_files(td)
```

# Acknowledgement

This project was funded in part by a United States Department of Agriculture—National Institute of Food and Agriculture (USDA—NIFA) Food Security Program Grant (Award Number 2016-68004-24769) and by United States Department of Agriculture (USDA) -Natural Resources Conservation Service (NRCS), Commodity Credit Corporation (CCC), Conservation Innovation Grants On-Farm Conservation Innovation Trials (Award Number USDA-NRCS-NHQ-CIGOFT-20-GEN0010750).
