---
title: "Trial Design Generation Report for _trial-name_"
output:
  bookdown::html_document2:
    toc: false
    number_sections: false
---

<style type="text/css">

body{ /* Normal  */
      font-size: 20px;
  }
td {  /* Table  */
  font-size: 16px;
}
h1.title {
  font-size: 56px;
}
h1 { /* Header 1 */
  font-size: 48px;
}
h2 { /* Header 2 */
    font-size: 36px;
}
h3 { /* Header 3 */
  font-size: 24px;
}
code.r{ /* Code block */
    font-size: 16px;
}
</style>

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  cache = FALSE,
  echo = TRUE,
  error = TRUE,
  warning = FALSE,
  cache.lazy = FALSE,
  message = FALSE,
  fig.width = 12, 
  fig.height = 12
)
```

```{r packages, message=FALSE, warning=FALSE, include=FALSE}
all_trial_info = readRDS(file.path("_temp-folder-here_", "all_trial_info.rds"))
machine_table = readRDS(file.path("_temp-folder-here_", "machine_table.rds"))

# === packages ===#
# library(bookdown)
# library(pracma)
# library(furrr)
# library(exactextractr)
# library(sp)
# library(sf)
# library(agricolae)
# library(lwgeom)
# library(measurements)
# library(stringr)
# library(tmap)
# library(raster)
# library(tidyverse)
# library(data.table)
# library(purrr)
# library(DT)
# library(geojsonsf)
# library(rmarkdown)
# library(knitr)
# library(parallel)
# library(jsonlite)
# library(smoother)
# library(zip)
```

```{r map-layout, message=FALSE, warning=FALSE, include=FALSE}
tm_layout_to_add <- tm_layout(
  legend.outside = "TRUE",
  frame = FALSE,
  legend.title.size = 2,
  legend.text.size = 1.5
)
```

## Description of Field Trial Design

Figure \@ref(fig:td-map) depicts the trial design made for the _trial-name_ trial.  Note that in the interior of the _field-size_-_land-unit_ field contains a grid of `r text_plot_num_length(all_trial_info, units)` Each plot is assigned one of `r text_rate_number(all_trial_info)` The variable rate `r trial_text_machinery_names_lower(machine_table)` `r text_plant_apply(machine_table)` `r trial_text_inputs(all_trial_info)` in each plot at a rate close to its targeted rate.  The number of plots assigned each targeted `r trial_text_inputs(all_trial_info)` rate is shown in Table \@ref(tab:rep-table).  Inside the circumference of the field there are _headland-size_-_length-unit_ headlands and _sideland-size_-_length-unit_ sidelands, which are not part of the field trial.

```{r td-map, echo=FALSE, fig.cap = "Trial Map", fig.height = 6, include = TRUE}

tmap_arrange(all_trial_info$map_design[[1]], all_trial_info$map_design[[2]], ncol = 2)

```

```{r rep-table, echo=FALSE, include = TRUE}
rep_table <- trial_data_r %>%
    rowwise() %>%
mutate(rep_table = list(
  trial_design %>% group_by(rate_tm) %>% 
  summarise(Replicates = n(),
            .groups = 'drop') %>%
  data.frame() %>%
  dplyr::select(rate_tm, Replicates) %>%
  dplyr::rename("Target Rate" = "rate_tm")))

knitr::kable(list(rep_table$rep_table[[1]], rep_table$rep_table[[2]]),
               caption = paste0("Replicate of each ", rep_table$input_name[[1]], " and ", rep_table$input_name[[2]], " Target Rate"),
               booktabs = TRUE,
               valign = 't')

```

## Steps in a Field Trial 

### Step 1. Apply Inputs Border Areas

Because areas near the border of the field are used to turn machinery around and may be exposed to different amounts of sun and wind than are areas in the field’s interior, data from border areas are not used in the experiment. Therefore, the first step in putting the trial design “in the ground” is to apply the input around border areas, in the same way the farmer would if not participating in the trial design. Figure \@ref(fig:headland-map) identifies the border areas for this field. 

```{r headland-map, echo = FALSE, fig.cap = "Map of Field Borders", fig.height = 6, include = TRUE}
all_trial_info$map_headlands[[1]]
```

### Step 2.  Center the `r trial_text_machinery_names_cap(machine_table)` on Machine’s AB-line Provided

One the trial design is made, its shapefile should be uploaded into the application monitor, as any other commercial site-specific input application prescription would be loaded.  Then the operator should center the machinery on the corresponding AB-line labeled “ab-lines-farmer-applicator” or “ab-lines-farmer-applicator” provided in the trial design files.  Figure \@ref(fig:machine-alignment) shows the `r trial_text_machinery_names_lower(machine_table)` inside a trial plot along with the respective AB-line. `r trial_text_ablines(machine_table)` This is the intended alignment given the `r trial_text_machine_sizes(machine_table, units)` and `r text_plot_width(all_trial_info, units)`. 

```{r machine-alignment, echo=FALSE, fig.cap="Alignment of Machinery inside a Trial Plot", fig.height=6, message=FALSE, warning=FALSE, include = TRUE}
ab_maps <- list()
poly_maps <- list()
label_maps <- list()
for (i in 1:(nrow(machine_table)-1)){
    ab_maps[[i]] <- paste0("machine_table$map_ab[[", i, "]]")
    poly_maps[[i]] <- paste0("machine_table$map_poly[[", i, "]]")
    label_maps[[i]] <- paste0("machine_table$map_label[[", i, "]]")
}

tm_shape(machine_table$trial_plot[[1]], bbox = st_bbox(machine_table$trial_plot[[1]])) +
  tm_borders(col = "black") +
  eval(parse(text = paste0(ab_maps, collapse = " + "))) +
  eval(parse(text = paste0(poly_maps, collapse = " + "))) +
  eval(parse(text = paste0(label_maps, collapse = " + "))) +
  tm_add_legend(type = "line",
                labels = c("Planter AB-line", "Applicator AB-line"),
                col = c("#009E73", "#0072B2"),
                lty = c("dashed", "solid"),
                title = "AB-Lines") +
  tm_add_legend(type = "symbol",
                labels = c("Planter", "Applicator"),
                col = c("#009E73", "#0072B2"),
                shape = 0,
                title = "Machine Polygons") +
  tm_add_legend(type = "symbol",
                labels = c("Trial Plot"),
                col = c("black"),
                shape = 0) +
  tm_layout(legend.outside = TRUE,
            frame = FALSE,
            bg.color = "grey85")
  

```

### Step 4. Uploading As-applied Data to Farmer’s DIFM Database 

After the applying the input to put the trial in the ground, the “as-applied” data should be downloaded from the application monitor and uploaded into the farmer’s DIFM database.   

### Step 5. Harvest 

When the trial design is created, the downloaded files also include a shapefile labeled "ab-lines-farmer-harvester", which the farmer should use to center the harvester at harvest time. Figure \@ref(fig:machine-alignment-harvester) shows the harvester and its AB-line along with all other machinery inside a trial plot. If the harvester is correctly centered, then the `r machine_table[nrow(machine_table), ]$number_in_plot` `r conv_unit(machine_table[nrow(machine_table), ]$width, "m", "ft")`-_length-unit_ harvester swaths will lie neatly within each `r conv_unit(td$input_plot_width[[1]], "m", "ft")`-_length-unit_ trial plot.   

```{r machine-alignment-harvester, echo=FALSE, fig.cap="Alignment of Machinery inside a Trial Plot", fig.height=6, message=FALSE, warning=FALSE, include = TRUE}
ab_maps <- list()
poly_maps <- list()
label_maps <- list()
for (i in 1:nrow(machine_table)){
    ab_maps[[i]] <- paste0("machine_table$map_ab[[", i, "]]")
    poly_maps[[i]] <- paste0("machine_table$map_poly[[", i, "]]")
    label_maps[[i]] <- paste0("machine_table$map_label[[", i, "]]")
}

tm_shape(machine_table$trial_plot[[1]], bbox = st_bbox(machine_table$trial_plot[[1]])) +
  tm_borders(col = "black") +
  eval(parse(text = paste0(ab_maps, collapse = " + "))) +
  eval(parse(text = paste0(poly_maps, collapse = " + "))) +
  eval(parse(text = paste0(label_maps, collapse = " + "))) +
  tm_add_legend(type = "line",
                labels = c("Harvester AB-line", "Planter AB-line", "Applicator AB-line"),
                col = c("#E69F00", "#009E73", "#0072B2"),
                lty = c("dashed", "dashed", "solid"),
                title = "AB-Lines") +
  tm_add_legend(type = "symbol",
                labels = c("Harvester", "Planter", "Applicator"),
                col = c("#E69F00", "#009E73", "#0072B2"),
                shape = 0,
                title = "Machine Polygons") +
  tm_add_legend(type = "symbol",
                labels = c("Trial Plot"),
                col = c("black"),
                shape = 0) +
  tm_layout(legend.outside = TRUE,
            frame = FALSE,
            bg.color = "grey85")
```

### Step 6.  Uploading Raw Yield Data into Farmer’s DIFM Database 

After the harvest, the raw yield data should be downloaded from the yield monitor and uploaded into the farmer’s DIFM database. 

### Step 7.  Uploading Field Characteristics Spatial Data into Farmer’s DIFM Database 

Once a boundary file is uploaded into the DIFM system, various spatial “characteristics” data will be automatically uploaded from public sources.  For example, data for virtually all agricultural fields in the USA can be downloaded from the USDA Natural Resources Conservation Service’s Web Soil Survey webpage (https://websoilsurvey.nrcs.usda.gov/app/) provides public access to the Soil Survey Geographic (SSURGO) Database, which contains soil maps.  Other data, such as elevation data and infrared satellite imagery are accessible through public sources. 

If you have any questions or would like to sign up as an official DIFM farmer or crop consultant, contact our DIFM Trial Supervisor Bob Dunker by email at redunker@illinois.edu.
