---
title: "Trial Design Options"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trial Design Options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 6,
  warning = FALSE,
  message = FALSE
)
```

The `ofpetrial` package offers various trial design types. This vignette provides examples of availabl design types. Let's first  prepare experiment plots to which we assign rates using various trial design options.

```{r, warning = FALSE, message = FALSE}
library(ofpetrial)

n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 30,
    section_num = 1,
    harvester_width = 20,
    headland_length = 30,
    side_length = 60
  )

exp_data <-
  make_exp_plots(
    input_plot_info = n_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )

viz(exp_data, type = "layout")
```

We will be assigning rates to the experimental plots using various trial design types below.

# Latin Square ("ls")

By default, here is what `design_type = "ls"` produces.

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "ls"
  )

td_ls_d <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td_ls_d)
```

However, you can customize the spatial pattern of input rate when `design_type = "ls"` using the `rank_seq_ws` and `rank_seq_as` options. To do so, it is important to understand how `plot_id` and `strip_id` are assigned to each of the plots. Here are their maps.

```{r, fig.height = 8}
ggpubr::ggarrange(
  viz(td_ls_d, type = "strip_id"),
  viz(td_ls_d, type = "plot_id"),
  ncol = 2
)
```

As you can see, `plot_id` is the unique numeric identifier assigned to each of the plots **within** a strip. So, there are multiple plots with the same `plot_id` values, but a combination of `strip_id` and `plot_id` uniquely identifies a plot.

The `rank_seq_ws` option specifies the order of rate rankings to be repetead **withing** a strip (this is why `_ws` at the end of the function). Suppose you have `rank_seq_ws = c(5, 4, 3, 2, 1)`. `5` refers to the 5th-ranked (highest) rate, which is 260 because we have `rates = c(100, 140, 180, 220, 260)` above. `1` refers to the first-ranked (lowest) rate, which is 100. Rates are assigned in this order to the plots within a strip. The `rank_seq_as` option specifies the order of the rate rankings of **the very first** plot of each strip **across** all the strips. So, for example, if `rank_seq_as = c(1, 4, 3, 2, 5)`, then the first plot (`plot_id == 1`) of the first strip (`strip_id == 1`) will be assigned rate rank of 1. The first plot of 5th strip ((`strip_id == 5`)) will be assigned rate rank of 5. This sequence will be repeated until the first plot of all the strips are assigned a rate rank. Now, for a given strip, rate ranks specified by `rank_seq_ws` will be repeated **starting from** the rate rank of the first plot. For example, the first plot of the 3rd strip has a rate rank of 3 (so, 180). This means that the code will go over the rest of the rate ranks in `rank_seq_ws` (2 and 1), and then go back to the beginning of `rank_seq_ws`, which is 5. So, for the third strip, the rate rank of its plots look like this.

```{r}
rank_seq_ws <- c(5, 4, 3, 2, 1)

data.frame(
  plot_id = 1:10,
  rate_rank = c(3, 2, 1, rank_seq_ws, rank_seq_ws[1:2])
)
```

Let's try few examples.

```{r}
n_rate_info$rank_seq_ws <- list(c(1, 2, 3, 4, 5))
n_rate_info$rank_seq_as <- list(c(1, 2, 3, 4, 5))

td_ls_1 <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td_ls_1, type = "rates")
```

```{r}
n_rate_info$rank_seq_ws <- list(c(5, 2, 4, 1, 3))
n_rate_info$rank_seq_as <- list(c(1, 5, 2, 4, 3))

td_ls_2 <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td_ls_2, type = "rates")
```

# Strip trial ("strip")

You can design a strip trial using `design_type = "strip"`. You can specify the spatial pattern of strip rates using the `rank_seq_as` option. The code below repeats 100 (rank 1), 280 (rank 4), 220 (rank 3), and 160 (rank 2). Since the strip trial has a single rate per strip, `rank_seq_ws` is not available unlike `design_type = "ls"`.

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 160, 220, 280),
    rank_seq_as = c(1, 4, 3, 2),
    design_type = "strip",
  )

td_strip <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td_strip, type = "rates")
```

For the strip trial, you can specify the full sequence unlike the other design options. We have a total of `r max(exp_data$exp_plots[[1]]$strip_id)` strips in this experiment.

```{r}
#--- total number of strips ---#
max(exp_data$exp_plots[[1]]$strip_id)
```

Let's provide a vector of length `r max(exp_data$exp_plots[[1]]$strip_id)` to `rank_seq_as`.

```{r}
n_rate_info$rank_seq_as <- list(c(1, 4, 3, 2, 2, 3, 4, 1, 3, 2, 1, 4, 2, 3, 1, 4, 4, 2, 3))

td_strip <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td_strip, type = "rates")
```

# Randomized Block ("rb")

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "rb",
  )

td_rb <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td_rb, type = "rates")
```

When `design_type = "rb"`, blocks are created internally when `assign_rates()` is run. Here is what blocks look like.

```{r}
add_blocks(td_rb) %>% viz(type = "block_id")
```

Since there are five distinctive rates, each block consists of five by five plots. In each of the block, the five rates are randomly assigned in a way that each of the rates appear exactly five times.

# Jump-conscious Latin Square ("jcls")

Jump-conscious Latin Square (`design_type = "jcls`) is merely a special type of Latin Square design. Indeed, you can create exactly the same design by specifying `rank_seq_as` and `rank_seq_ws` appropriately. JCLS ensures that the jumps in rate from plot to plot in the direction of the applicator/planter less than equal to rate rank of 2. For exmaple, input rate never goes from 100 to 220 or 260 in JCLS.  

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "jcls",
  )

td_jcls <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td_jcls, type = "rates")
```

# Extra Jump-conscious Alternate ("ejca")

This design alternate high-rate strip and low-rate strip, thus avoiding sudden changes in input rates so that machines can handle them.

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260, 300),
    design_type = "ejca",
  )

td_ejca <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td_ejca, type = "rates")
```

# Sparse ("sparse")

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260, 300),
    design_type = "sparse",
  )

td_sparse <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )
```

```{r}
viz(td_sparse, type = "rates")
```

