---
title: "Change Trial Rates Manually"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{chante-rates-manually}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 6
)
```

```{r setup}
library(ofpetrial)
library(ggplot2)
library(dplyr)
```

Suppose you have created a trial design is made using `assign_rates()` to experiment plots made by `make_exp_plots()`, but you are not quite satisfied with it and would like to change rates here and there. You can easily do so manually. This vignette demonstrates such operations. 

# Data Preparation 

Let's first create a trial design for a single input case.

```{r}
n_plot_info <-
  prep_plot_fs(
    input_name = "NH3",
    machine_width = 30,
    section_num = 1,
    harvester_width = 20,
    headland_length = 30,
    side_length = 60
  )

exp_data <-
  make_exp_plots(
    input_plot_info = n_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )

n_rate_info <-
  prep_rates_s(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "ls",
    rank_seq_ws = c(1, 2, 3, 4, 5),
    rank_seq_as = c(1, 2, 3, 4, 5)
  )

td <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )
```

Here is what the trial design looks like.

```{r}
viz(td, type = "rates")
```

# Changing rates 

It is important to be aware that every single plot is assigned a withint-strip plot id and a strip id when they were created using `make_exp_plots()`. 

```{r}
td$trial_design[[1]]
```

The figure below shows the strip id associated with each plot.
```{r, fig.height = 9}
viz(td, type = "strip_id")
```

The figure below shows the plot id associated with each plot.
```{r, fig.height = 9}
viz(td, type = "plot_id")
```

As you can see, `plot_id` is the unique numeric identifier assigned to each of the plots **within** a strip. So, there are multiple plots with the same `plot_id` values, but a combination of `strip_id` and `plot_id` uniquely identifies a plot.

## By plot

You can use `change_rate_by_plot()` to change rates. The code below change the rate asscociated the plot with `strip_id = 1` and `plot_id = 1` (left bottom cell) to 0.

```{r}
modified_td <- 
  change_rate_by_plot(
    td,
    input_name = "NH3",
    strip_ids = 1,
    plot_ids = 1,
    new_rate = 0
  )
```

```{r}
viz(modified_td, type = "rates", abline = FALSE)
```

You can change rates of multiple plots with the same `plot_id` across multiple `strip_id`s.

```{r}
change_rate_by_plot(
  td,
  input_name = "NH3",
  strip_ids = 1:10,
  plot_ids = 10,
  new_rate = 0
) %>%
viz(abline = FALSE)
```

When you give vectors of values to `stip_ids` and `plot_ids`, the plots with all the possible combinations of `strip_id`-`plot_id` are going to have a new rate.

```{r}
change_rate_by_plot(
  td,
  input_name = "NH3",
  strip_ids = 1:10,
  plot_ids = c(1, 5, 9, 14, 19, 24),
  new_rate = 0
) %>%
viz(abline = FALSE)
```

## By strip

`change_rate_by_strip()` lets you change the rate of all the plot in the strips you specified. When you supply vectoros of equal length to `strip_ids` and `new_rates`, then **n**th element of `strip_ids` is matched with **n**th element of `new_rates`.

```{r}
change_rate_by_strip(
  td,
  input_name = "NH3",
  strip_ids = c(1, 4, 7, 10, 13, 16, 19),
  new_rates = seq(0, 270, length = 7)
) %>%
viz(abline = FALSE)
```

If you supply a single value to `new_rates` while supplying more than one values to `strip_ids`, the single rate will be assigned to all the strips specified in `strip_ids`.

```{r}
change_rate_by_strip(
  td,
  input_name = "NH3",
  strip_ids = c(1, 4, 7, 10, 13, 16, 19),
  new_rates = 0
) %>%
viz(abline = FALSE)
```

## By block

By default, experiment plots made using `make_exp_plots()` are not assigned block IDs. You can do so using `add_blocks()`.

```{r, fig.height = 9}
td_with_blocks <- add_blocks(td)

viz(td_with_blocks, type = "block_id")
```