---
title: "More Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Work Flow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 6
)
```

```{r setup}
library(ofpetrial)
```

# Single-input Nitrogen Experiment

Create plot information.

```{r}
n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 30,
    section_num = 1,
    harvester_width = 30,
    plot_width = 30
  )
```

Create exerpimental plots based on the plot information.

```{r}
exp_data <-
  make_exp_plots(
    input_plot_info = n_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )
```

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "ls",
  )

trial_design <- assign_rates(exp_data, rate_info = n_rate_info)
```

```{r, dpi = 600}
viz(trial_design)
```

# Two-input Nitrogen and Seed Experiment

```{r}
n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 30,
    section_num = 1,
    harvester_width = 30,
    plot_width = 30
  )

n_plot_info

seed_plot_info <-
  prep_plot(
    input_name = "seed",
    unit_system = "imperial",
    machine_width = 60,
    section_num = 24,
    harvester_width = 30,
    plot_width = 30
  )

seed_plot_info
```

```{r}
input_plot_info <- list(n_plot_info, seed_plot_info)

exp_data <-
  make_exp_plots(
    input_plot_info = input_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )
```

```{r, dpi = 600}
viz(exp_data, type = "layout", abline = TRUE)
```

```{r}
#!===========================================================
# ! Assign rates
# !===========================================================
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "ls",
  )

n_rate_info

seed_rate_info <-
  prep_rate(
    plot_info = seed_plot_info,
    gc_rate = 32000,
    unit = "seed",
    min_rate = 16000,
    max_rate = 40000,
    num_rates = 5,
    design_type = "ls"
  )

seed_rate_info
```

```{r}
trial_design <- assign_rates(exp_data, rate_info = list(n_rate_info, seed_rate_info))
```

```{r, dpi = 600}
viz(trial_design)
```

# Cover crop experiment 

# Precision Nitrogen Project (PNP) at University of Nebraska Lincoln

Three types

Two strips 
+ Grower's business-as-usual flat-rate (say, 200 lbs/acre of N)
+ Model-based variale-rate (how does this work?)
+ N-reponse block

```{r}
n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 60,
    section_num = 1,
    harvester_width = 30,
    headland_length = 30,
    plot_width = 60,
    min_plot_length = 200,
    max_plot_length = 300
  )

exp_data <-
  make_exp_plots(
    input_plot_info = n_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )

viz(exp_data, type = "layout")
```

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 240,
    unit = "lb",
    rates = 240,
    design_type = "str"
  )

td <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td, type = "rates")
```

```{r}
td <-
  change_rates(
    td,
    strip_ids = c(3, 6),
    new_rates = 180
  )

viz(td, type = "rates")
```

```{r}
td <-
  td %>%
  change_rates(
    strip_ids = 8:12,
    plot_ids = 5:9,
    new_rates = c(100, 140, 180, 220, 260),
    rate_by = "strip"
  ) %>%
  change_rates(
    strip_ids = 8:12,
    plot_ids = 15:19,
    new_rates = c(140, 260, 100, 180, 220),
    rate_by = "strip"
  ) %>%
  change_rates(
    strip_ids = 8:12,
    plot_ids = 25:29,
    new_rates = c(220, 100, 260, 220, 180),
    rate_by = "strip"
  )

viz(td, type = "rates")
```

```{r}
td <-
  change_rates(
    td,
    strip_ids = c(15, 16),
    new_rates = 180
  )

viz(td, type = "rates")
```

