---
title: "Modify or create trial design maps from scratch"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modify or create trial design maps from scratch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

By default, `viz()` uses the `Greens` palette, but you may want to change the color scheme. Or you may want to change some other parts of the maps. You can do this realtively easily with knowledge of  `ggplot2` package.

Let's create a trial design first.

```{r}
#--- packages ---#
library(ofpetrial)
library(ggplot2)

n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 30,
    section_num = 1,
    harvester_width = 30,
    plot_width = 30
  )

exp_data <-
  make_exp_plots(
    input_plot_info = n_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )

n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "ls",
    rank_seq_ws = c(5, 4, 3, 2, 1)
  )

td <- assign_rates(exp_data, rate_info = n_rate_info)
```

```{r}
#--- create a map using viz() ---#
viz(td)
```

# Modify the map created by `viz()`

While running `viz()` displays a plot, you can save it to a new object like below. 


```{r}
#--- save it to a new object ---#
viz_td <- viz(td)

#--- check its class ---#
class(viz_td)
```

As you can see, it is a `ggplot` object. This means that you can for example change the fill color scheme to viridis like below (along with change the title):

```{r}
viz_td +
  scale_fill_viridis_d(name = "NH3 (lb/ac)| N Equivalent (lb/ac)") +
  ggtitle("New Title")
```

# Create a trial design map from scratch

The `viz()` uses trail design `sf` object stored in `td` (trial design created using `assign_rates()`), which is named `trial_design`. You can access it like below:

```{r}
#--- extract the trial design sf ---#
td_for_plot <- td$trial_design[[1]]

#--- check the class ---#
class(td_for_plot)
```

So, you can use the `ggplot2` package (or other pacakges like `tmap`) to create maps yourself.

```{r}
ggplot(data = td_for_plot) +
  geom_sf(aes(fill = factor(rate)))
```

This is not a `ggplot2` guideline. So, we stop here. But, it should be clear at this point that you can further modify the map to your liking using `ggplot2`.
