---
title: "Change Trial Rates Manually"
output: rmarkdown::html_vignette
bibliography: reference.bib
vignette: >
  %\VignetteIndexEntry{Change Trial Rates Manually}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 6
)
```

```{r setup, include = FALSE}
library(ofpetrial)
library(ggplot2)
library(dplyr)
```

```{r setup-show, eval = FALSE}
library(ofpetrial)
library(ggplot2)
library(dplyr)
```

Suppose you have created a trial design applying `assign_rates()` to experiment plots made by `make_exp_plots()`, but you are not quite satisfied with it and would like to change rates here and there. You can easily do so manually using the `change_rates()` function. This vignette demonstrates such operations. 

# Data Preparation 

Let's first create a trial design for a single input case.

```{r}
n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 30,
    section_num = 1,
    harvester_width = 20,
    headland_length = 30,
    side_length = 60
  )

exp_data <-
  make_exp_plots(
    input_plot_info = n_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )

n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 180,
    unit = "lb",
    rates = c(100, 140, 180, 220, 260),
    design_type = "ls",
    rank_seq_ws = c(1, 2, 3, 4, 5),
    rank_seq_as = c(1, 2, 3, 4, 5)
  )

td <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )
```

Here is what the trial design looks like.

```{r}
viz(td, type = "rates")
```

# Changing rates 

It is important to be aware that every single plot is assigned a withint-strip plot id and a strip id when they were created using `make_exp_plots()`. 

```{r}
td$trial_design[[1]]
```

The figure below shows the strip id associated with each plot.
```{r, fig.height = 9}
viz(td, type = "strip_id")
```

The figure below shows the plot id associated with each plot.
```{r, fig.height = 9}
viz(td, type = "plot_id")
```

As you can see, `plot_id` is the unique numeric identifier assigned to each of the plots **within** a strip. So, there are multiple plots with the same `plot_id` values, but a combination of `strip_id` and `plot_id` uniquely identifies a plot.

## Re-assign a single rate to all the specified plots

You can use `change_rates()` to change rates. The code below change the rate asscociated the plot with `strip_id = 1` and `plot_id = 1` (left bottom cell) to 0.

```{r}
modified_td <-
  change_rates(
    td = td,
    input_name = "NH3",
    strip_ids = 1,
    plot_ids = 1,
    new_rates = 0
  )
```

```{r}
viz(modified_td, type = "rates")
```

You can change rates of multiple plots with the same `plot_id` across multiple `strip_id`s.

```{r}
change_rates(
  td,
  input_name = "NH3",
  strip_ids = 1:10,
  plot_ids = 10,
  new_rates = 0
) %>%
  viz(abline = FALSE)
```

When you give vectors of values to `stip_ids` and `plot_ids`, the plots with all the possible combinations of `strip_id`-`plot_id` are going to have a new rate.

```{r}
change_rates(
  td,
  input_name = "NH3",
  strip_ids = 1:10,
  plot_ids = c(1, 5, 9, 14, 19, 24),
  new_rates = 0
) %>%
  viz(abline = FALSE)
```

If you do not provide any `plot_ids`, then all the plots in the strips specified by `strip_ids` will be altered. 

```{r}
change_rates(
  td,
  input_name = "NH3",
  strip_ids = c(1, 4, 7, 10, 13, 16),
  new_rates = 0
) %>%
  viz(abline = FALSE)
```

## Re-assign multiple rates by strip

You can also change the rate of multiple plots across multiple strips by strip using the `rate_by = "strip"` option. In this case, **n**th element of `new_rates` will be assigned to **n**th strip in `strip_ids`.

```{r}
change_rates(
  td,
  input_name = "NH3",
  strip_ids = 2:7,
  plot_ids = 3:8,
  new_rates = c(0, 300, 0, 300, 0, 300),
  rate_by = "strip"
) %>%
  viz(abline = FALSE)
```

If you leave `plot_ids` unspecified, all the plots in the strips specified by `strip_ids` will be assigned new rates. That is, rates are re-assigned by strip. 

```{r}
change_rates(
  td,
  input_name = "NH3",
  strip_ids = 2:7,
  new_rates = c(0, 300, 150, 200, 100, 300),
  rate_by = "strip"
) %>%
  viz(abline = FALSE)
```

The `rate_by = "strip"` can be useful when you want to "remove" some plots from the experiment. For example, take a look at this trial design.

```{r}
data(td_curved)
viz(td_curved)
```

Notice that there is a three-plot strip at the east end of the field. Suppose, you would like to "remove" it from experiments (there is nothing wrong with keeping them as part of the experiment). You can do so by assining the rate for the non-experimental part of field (which is `r td_curved$gc_rate`) to the strip. 

```{r}
change_rates(
  td_curved,
  input_name = "NH3",
  strip_ids = 31,
  new_rates = 180,
  rate_by = "strip"
) %>%
  viz(abline = FALSE)
```

## Re-assign multiple rates by plot

You can assign new rates plot by plot at one time using the `rate_by = "plot"` option. For this option, you are asked to provide a matrix of rates where rows and columns of the matrix represents `plot_id`s and `strip_id`s, respectively. For example, suppose you have `plot_ids = c(2, 4, 6)` and `strip_ids = c(1, 3, 5, 7)`. Further suppose, you have specified `new_rates` like below.

```{r}
(
  new_rates <- matrix(1:12 * 20, nrow = 3, ncol = 4)
)
```

Then, the plot with `plot_id` = 2 (1st element of `plot_ids`) and `strip_id` = 1 (1st element of `strip_ids`) will get the value stored in `new_rates[1, 1]` (1st row and 1st column of the matrix of rates).

The code below assign randomized rates to the 5-plot by 5-plot block at the left lower corner of the field. 

```{r}
new_rates_mat <- replicate(5, sample(c(100, 140, 180, 220, 260), 5))

change_rates(
  td,
  input_name = "NH3",
  strip_ids = 1:5,
  plot_ids = 1:5,
  new_rates = new_rates_mat,
  rate_by = "plot"
) %>%
  viz(abline = FALSE)
```

If you would like, you could repeat this to other blocks of plots to generate a randomized block design. Of course, it is much easier to just use the "rb" option in `assign_rates()`.

# Application: Precision Nitrogen Project (PNP) at University of Nebraska Lincoln

Here, we apply what we have learned so far to create a trial design like the ones used in @puntel2024leveraging.

```{r}
n_plot_info <-
  prep_plot(
    input_name = "NH3",
    unit_system = "imperial",
    machine_width = 60,
    section_num = 1,
    harvester_width = 30,
    headland_length = 30,
    plot_width = 60,
    min_plot_length = 200,
    max_plot_length = 300
  )

exp_data <-
  make_exp_plots(
    input_plot_info = n_plot_info,
    boundary_data = system.file("extdata", "boundary-simple1.shp", package = "ofpetrial"),
    abline_data = system.file("extdata", "ab-line-simple1.shp", package = "ofpetrial"),
    abline_type = "free"
  )

viz(exp_data, type = "layout")
```

```{r}
n_rate_info <-
  prep_rate(
    plot_info = n_plot_info,
    gc_rate = 240,
    unit = "lb",
    rates = 240,
    design_type = "str"
  )

td <-
  assign_rates(
    exp_data = exp_data,
    rate_info = n_rate_info
  )

viz(td, type = "rates")
```

```{r}
td <-
  change_rates(
    td,
    strip_ids = c(3, 6),
    new_rates = 180
  )

viz(td, type = "rates")
```

```{r}
td <-
  td %>%
  change_rates(
    strip_ids = 8:12,
    plot_ids = 5:9,
    new_rates = c(100, 140, 180, 220, 260),
    rate_by = "strip"
  ) %>%
  change_rates(
    strip_ids = 8:12,
    plot_ids = 15:19,
    new_rates = c(140, 260, 100, 180, 220),
    rate_by = "strip"
  ) %>%
  change_rates(
    strip_ids = 8:12,
    plot_ids = 25:29,
    new_rates = c(220, 100, 260, 220, 180),
    rate_by = "strip"
  )

viz(td, type = "rates")
```

```{r}
td <-
  change_rates(
    td,
    strip_ids = c(15, 16),
    new_rates = 180
  )

viz(td, type = "rates")
```

# References